Algebraic filtering of surfaces from medical images with Julia

### 2020-04-26 src/background.tex
submission

\subsection{Representation Schemes}\label{sec:schemes}

A \emph{representation scheme} for solid modeling is a mapping between a space of mathematical models and  space of symbolic representations like generated by a formal grammar.
Solid pointsets (i.e., `$r$-sets') are defined~\cite{Requicha:1980:RRS:356827.356833} as compact (bounded and closed) regular and semianalytic subsets of the $d$-space. A large number of representation schemes were defined in the past forty years, including the two main classes of (a) \emph{boundary representations} (`$B$-reps'), where the solid model is represented through a representation of its boundary elements, i.e.~faces, edges and vertices, and (b) \emph{decompositive/enumerative representations}, that are a decomposition of either the object or the embedding space, respectively, into a well-defined \emph{cellular complex}. In particular, a boundary representation provides a cellular decomposition of the object's boundary into \emph{cells} of dimension zero (vertices), one (edges), and two (faces). Medical imaging can be classified as the \emph{enumerative representation} of cellular decompositions of organs and tissues of interest, in particular, as subsets \emph{of 3D volume elements} (voxels) from the 3D image.

@@

The \emph{Linear Algebraic Representation} (\textsc{lar}), introduced in~\cite{Dicarlo:2014:TNL:2543138.2543294}, aims to represent the \emph{chain complex}~\cite{TSAS} generated by a piecewise-linear \emph{geometric complex} embedded either in 2D or in 3D. In a few words, it gets a minimal characterization of geometry and topology of a cellular complex, i.e.~the embedding mapping $\mu : C_0 \to \E^d$ of 0-cells (vertices), as well a description of $(d-1)$-cells as subsets of vertices, and is able to return the whole chain complex 
\begin{equation}
C_\bullet = (C_p, \partial_p) := 
C_3 
\substack{
\delta_2 \\
\longleftarrow \\
\longrightarrow \\
\partial_3 
}
C_2 
\substack{
\delta_1 \\
\longleftarrow \\
\longrightarrow \\
\partial_2 
}
C_1  
\substack{
\delta_0 \\
\longleftarrow  \\
\longrightarrow \\
\partial_1 
}
C_0 .
\end{equation}

@@

It is possible to show that all the interesting relations of incidence/adjacency between cells of different dimensions can be both computed and efficiently queried by pairwise computing some matrix products, with one of terms possibly transposed, using only the two boundary and coboundary operator matrices $[\partial_p]$ and $[\delta_p]$, and where $[\delta_p] = [\partial_p^\top]$. We~may also show that such matrices are \emph{very sparse}, with their sparseness growing rapidly with the dimensions (see Section~\ref{sec:block}). The pattern of non-zeros in matrix $[\partial_3]$ corresponding to a brick of shape $(4,4,4)$ is given in Fig.~\ref{fig:boundary_matrix_4x4x4}.

\begin{figure}[tbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.75\linewidth]{figs/boundary_matrix_4x4x4.pdf} 
   \caption{
   The binary image of \emph{sparse} coboundary matrix  $\left[\delta_2\right] = \left[\partial_3\right]^t : C_2 \to C_3$,
%   The binary image of the coboundary operator  $\delta_2 = \partial_3^\top : C_2 \to C_3$, 
   built for a small volumetric data (or a brick) with shape $(4,4,4)$. Note that the number of rows equates the size $4\times 4\times 4 = 64$ of the voxel set; the number of columns is $d\,n\,(1+n)^{d-1} = 3\times 4\times 25 = 300$. Of course, the number of non-zeros per row (cardinality of the facet set of a single voxel) is six, whereas the number of non-zeros per column is two, but on boundary facets.}
   \label{fig:boundary_matrix_4x4x4}
\end{figure}

>>>>

\subsection{Representation Scheme}\label{sec:schemes}

A \emph{representation scheme} for solid modeling is a mapping between a space of mathematical models and  space of symbolic representations like generated by a formal grammar.
Solid pointsets (i.e., `$r$-sets') are defined~\cite{Requicha:1980:RRS:356827.356833} as compact (bounded and closed) regular and semianalytic subsets of the $d$-space. A large number of representation schemes were defined in the past forty years, including the two main classes of (a) \emph{boundary representations} (`$B$-reps'), where the solid model is represented through a representation of its boundary elements, i.e.~faces, edges and vertices, and (b) \emph{decompositive/enumerative representations} \cite{Requicha:1980:RRS:356827.356833}, that are a decomposition of either the object or the embedding space, respectively, into a well-defined \emph{cellular complex}. In particular, a boundary representation provides a cellular decomposition of the object's boundary into \emph{cells} of dimension zero (vertices), one (edges), and two (faces). Medical imaging can be classified as the \emph{enumerative representation} of cellular decompositions of organs and tissues of interest \cite{Paoluzzi2016}, in particular, as subsets \emph{of 3D volume elements} (voxels) from the 3D image.

@@

The \emph{Linear Algebraic Representation} (\textsc{lar}), introduced in~\cite{Dicarlo:2014:TNL:2543138.2543294}, aims to represent the \emph{chain complex}~\cite{TSAS,DiCarlo2009} generated by a piecewise-linear \emph{geometric complex} embedded either in 2D or in 3D. In a few words, it gets a minimal characterization of geometry and topology of a cellular complex, i.e.~the embedding mapping $\mu : C_0 \to \E^d$ of 0-cells (vertices), as well a description of $(d-1)$-cells as subsets of vertices, and is able to return the whole chain complex 
\begin{equation}
C_\bullet = (C_p, \partial_p) := 
C_3 
\substack{
\delta_2 \\
\longleftarrow \\
\longrightarrow \\
\partial_3 
}
C_2 
\substack{
\delta_1 \\
\longleftarrow \\
\longrightarrow \\
\partial_2 
}
C_1  
\substack{
\delta_0 \\
\longleftarrow  \\
\longrightarrow \\
\partial_1 
}
C_0 .
\end{equation}

@@

It is possible to show that all the interesting relations of incidence/adjacency between cells of different dimensions can be both computed and efficiently queried by pairwise computing some matrix products, with one of terms possibly transposed, using only the two boundary and coboundary operator matrices $[\partial_p]$ and $[\delta_p]$, and where $[\delta_p] = [\partial_p^\top]$. We~may also show that such matrices are \emph{very sparse}, with their sparseness growing rapidly with the dimensions (see Section~\ref{sec:brick}). The pattern of non-zeros in matrix $[\partial_3]$ corresponding to a brick of shape $(4,4,4)$ is given in Fig.~\ref{fig:boundary_matrix_4x4x4}.

\begin{figure}[tbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.75\linewidth]{src/figs/boundary_matrix_4x4x4_desc.png} 
   \caption{
   The binary image of \emph{sparse} coboundary matrix  $\left[\delta_2\right] = \left[\partial_3\right]^t : C_2 \to C_3$,
%   The binary image of the coboundary operator  $\delta_2 = \partial_3^\top : C_2 \to C_3$, 
   built for a small volumetric data (or a brick) with shape $(4,4,4)$. Note that the number of rows equates the size $4\times 4\times 4 = 64$ of the voxel set; the number of columns is $d\,n\,(1+n)^{d-1} = 3\times 4\times 25 = 300$. Of course, the number of non-zeros per row (cardinality of the facet set of a single voxel) is six, whereas the number of non-zeros per column is two, but on boundary facets.}
   \label{fig:boundary_matrix_4x4x4}
\end{figure}

### 2020-04-26 src/conclusion.tex
submission

% TODO from abstract
We introduced a Julia implementation of an algebraic filter to extract from 3D medical images the
boundary surface of some specific image segment, described as a 3-chain of voxels. Translations from
Cartesian indices of cells to linearized indices, the computation of the sparse boundary matrices, and the
sparse matrix-vector multiplication are the main computational kernels of this approach. We may show
a good speed-up over marching-cubes algorithms. The existing implementation employs Julia's channels
for multiprocessing. Currently, the computational pipeline is being strongly improved to gain a greater
speed-up using native Julia implementation \texttt{CUDA.jl} of Nvidia programming platform 
% TODO cite Besard2017 [7]
, and the Julia's
\texttt{SuiteSparseGraphBLAS.jl} framework 
% TODO cite BULUK 2017 [8] 
for graph algorithms with the language of linear algebra. In
particular, we are extending its use pattern in order to work with general cellular complexes.

>>>>

% % TODO from abstract
% We introduced a Julia implementation of an algebraic filter to extract from 3D medical images the
% boundary surface of some specific image segment, described as a 3-chain of voxels. Translations from
% Cartesian indices of cells to linearized indices, the computation of the sparse boundary matrices, and the
% sparse matrix-vector multiplication are the main computational kernels of this approach. We may show
% a good speed-up over marching-cubes algorithms. The existing implementation employs Julia's channels
% for multiprocessing. Currently, the computational pipeline is being strongly improved to gain a greater
% speed-up using native Julia implementation \texttt{CUDA.jl} of Nvidia programming platform 
% % TODO cite Besard2017 [7]
% , and the Julia's
% \texttt{SuiteSparseGraphBLAS.jl} framework 
% % TODO cite BULUK 2017 [8] 
% for graph algorithms with the language of linear algebra. In
% particular, we are extending its use pattern in order to work with general cellular complexes.

We introduced a Julia implementation of an algebraic filter to extract from 3D medical images the
boundary surface of some specific image segment, described as a 3-chain of voxels. Translations from
Cartesian indices of cells to linearized indices, the computation of the sparse boundary matrices, and the
sparse matrix-vector multiplication are the main computational kernels of this approach.

The implementation of the \textsc{lar-surf} filter is available in the open-source repository and it can be installed using standard Julia package manager \cite{larsurf-github}.

% NEW next two sentences are new
We showed a good speed-up over marching-cubes algorithms. 
The existing implementation employs Julia's channels
for multiprocessing. 
Our performance experiment showed an optimal size of the brick size.
Parallelization makes a large portion of spared computational cost. 
%
Moreover, we expect additional improvement in the future because our approach is appropriate  for SIMD (Single Instruction, Multiple Data) hybrid architectures of CPUs and GPUs, since only the initial block setup of boundary matrix and image slices, as well the final collection of computed surface portions, require inter-process communication.

\sloppy{
Currently, the computational pipeline is being strongly improved to gain a greater
speed-up using native Julia implementation \texttt{CUDA.jl} of Nvidia programming platform 
% TODO cite Besard2017 [7]
% \cite{Besard:2017}
\cite{Besard2019}
, and Julia's
\texttt{SuiteSparseGraphBLAS.jl} framework 
\cite{Buluc2017}
% TODO cite BULUK 2017 [8] 
for graph algorithms with the language of linear algebra. In
particular, we are extending its use pattern in order to work with general cellular complexes.
}

### 2020-04-26 src/design.tex
submission

\section{Block-parametric design}\label{sec:filter}

\subsection{Block decomposition}\label{sec:block-decomposition}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.5\linewidth]{figs/blocks} 
   \caption{A possible block partitioning of a radiologic image. The evidenced 2D block, of size $n^d=64^2$, is sliced by $\B([2,1,64]) = Image([128:172],[64:128)]$}
   \label{fig:blocks}
\end{figure}

Figure~\ref{fig:blocks} shows the block decomposition in a 2D image, with positive integers $(u,v)$ giving the lateral sizes of image. Note that block sides do not necessarily correspond to image edges.

\subsection{Block operator }\label{sec:block}

\subsubsection*{Chain coordinates }\label{sec:chain-coords}
We are going to treat each image block independently from each other. Hence we map each image block $\B(i,j,k,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n\times n$, using coordinate vectors $c\in \mathbf{2}^{n^d} := \{0,1\}^{n^d}$, where the basis element $c \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

Therefore, each pixel (or voxel) in a block image will be seen as a basis binary vector in $C_2$, and each subset of image elements, as the corresponding binary vector in $C_2$, with many ones as the cardinality of the subset.

\subsubsection*{Boundary operator }\label{sec:boundary-operator}
For a fixed block size $n$, the boundary operator $\partial_d : C_d\to C_{d-1}$, with $d\in\{2,3\}$, will be constructed once and for all using the algorithm given in~
\cite{TSAS}, 
% \cite{DiCarlo2014}, 
and inlined in the generated boundary extraction code.

@@

The number of rows of $[\partial_d]$ equates the dimension of the linear space $C_{d-1}$, i.e.~the number of $(d-1)$-cells---elementary $(d-1)$-chains---in the cellular partition of the image. To compute their number, we act in two steps. (a) First we map one-to-one the $n^d$ $d$-cells with $d$ adjacent $(d-1)$-cells, so getting $d\,n^d$ distinct basis elements of $C_{d-1}$. (b) Then we complete this bases by adjoining $n^{d-1}$ boundary elements for each of the $d$ dimensions of the image, so providing further $d\,n^{d-1}$ basis elements for $C_{d-1}$. The dimension of $C_{d-1}$, and therefore the number of rows of $[\partial_d]$ matrix is $d\,(n^{d-1}+n^{d}) = d\,n\,(1+n)^{d-1}$. The number of column equates the number of basis elements of $C_d$, i.e.~the number $n^d$ of block elements.

@@

As we have seen, we have $2d$ non-zero elements for each column of $[\partial_d]$, so that their total number is $2d\,n^d$. The number of matrix element is $d\,n\,(1+n)^{d-1} \times n^d$, giving a ratio of 
\[
\frac{\mbox{non-zero\ elements}}{\mbox{total\ elements}} = 
\frac{2d\times n^d}{d\,n\,(1+n)^{d-1} \times n^d} =
\frac{2}{n+n^d}
\]
Using sparse matrices in CSC (Compressed Sparse Column) format we get a storage size:
\[
mem([\partial_d]_{n^d}) = 2\times \#\mbox{nzero} + \#\mbox{columns} = 2\times 2d\,n^d + n^d = (4d+1)n^d.
\]
In conclusion, for block size $n=64$, the matrix $[\partial_d]$ requires for 2D images $9\times 64^2=36,864$ memory elements, and for 3D images $13\times 64^3=3,407,872$ memory elements. Counting the bytes for the standard implementation of a sparse binary matrix (1 byte for values and 8 bytes for indices) we get $(18d+8)n^d$ bytes, giving $176$\,KB for 2D and $15.872$\,MB for 3D.

% TODO fix the name
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.4\linewidth]{figs/grid.pdf} 
   \caption{Faces on the brick boundary}
   \label{fig:blockboundary}
\end{figure}

\subsection{Block boundary mapping}\label{sec:block-mapping}

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. Our aim is to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the block, in order to decompose the input $Image(u,v,w)$ into a fair number of blocks
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n} \simeq \frac{uvw}{n^3}.
\] 
Then, we consider each segment portion $c_{i,j,k} = S\cap \B(i,j,k,n)$ and compute its local coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then, assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$, with $d=3$. Finally, compute a matrix $\T{\textbf{B}}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{\textbf{B}} = [\partial_3(n)]\, \T{S}.
\]
where the boundary matrix has dimension $n^d \times dn(n+1)^{d-1}$.
Of course, the $\T{\textbf{B}}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S \cap \B(i,j,k,n)$, and the number of matrix rows is the dimension $n^d$ of the linear space $C_2$ of 2-chains.

@@

\subsection{Block-level parallelism}\label{sec:block-parallelism}
In the computational pipeline introduced in this paper, several steps can be efficiently performed in parallel at the image-block level, depending on the embarrassingly data-parallel nature of the problem. In particular, little effort is needed to separate the problem into a several of parallel tasks $S_{i,j,k}$, using multiarray slicing. The granularity of parallelism, depending on the block size $n$, is further enforced by the computation of a single boundary matrix $[\partial_d(n)]$ depending on $n$, so that the initial communication cost of broadcasting the matrix to nodes can be carefully controlled, and finely tuned depending on the system architecture. The whole approach is appropriate  for SIMD (Single Instruction, Multiple Data) hybrid architectures of CPUs and GPUs, since only the initial block setup of boundary matrix and image slices, as well the final collection of computed surface portions, require inter-process communication.

>>>>

\section{Brick-parametric design}\label{sec:filter}

\subsection{Brick decomposition}\label{sec:brick-decomposition}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
    \centering
    \begin{subfigure}{0.49\textwidth}
       \centering
       \includegraphics[width=0.70\linewidth]{src/figs/blocks.png} 
       \subcaption{A possible brick partitioning of a radiologic image. The evidenced 2D brick, of size $n^d=64^2$, is sliced by $\B([2,1,64]) = Image([128:172],[64:128)]$}
       \label{fig:bricks}
    \end{subfigure} 
    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics[width=0.99\linewidth]{figs/grid.pdf} 
        \subcaption{Faces on the brick boundary}
        \label{fig:brickboundary}
    \end{subfigure} 
    \caption{Brick decomposition}
\end{figure}

Figure~\ref{fig:bricks} shows the brick decomposition in a 2D image, with positive integers $(u,v)$ giving the lateral sizes of image. Note that brick sides do not necessarily correspond to image edges.

\subsection{Brick operator }\label{sec:brick}

\subsubsection*{Chain coordinates }\label{sec:chain-coords}
We are going to treat each image brick independently from each other. Hence we map each image brick $\B(i,j,k,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n\times n$, using coordinate vectors $c\in \mathbf{2}^{n^d} := \{0,1\}^{n^d}$, where the basis element $c \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

Therefore, each pixel (or voxel) in a brick image will be seen as a basis binary vector in $C_2$, and each subset of image elements, as the corresponding binary vector in $C_2$, with many ones as the cardinality of the subset.

\subsubsection*{Boundary operator }\label{sec:boundary-operator}
For a fixed brick size $n$, the boundary operator $\partial_d : C_d\to C_{d-1}$, with $d\in\{2,3\}$, will be constructed once and for all using the algorithm given in~
\cite{TSAS}, 
% \cite{DiCarlo2014}, 
and inlined in the generated boundary extraction code.

@@

The number of rows of $[\partial_d]$ equates the dimension of the linear space $C_{d-1}$, i.e.~the number of $(d-1)$-cells---elementary $(d-1)$-chains---in the cellular partition of the image. To compute their number, we act in two steps. (a) First we map one-to-one the $n^d$ $d$-cells with $d$ adjacent $(d-1)$-cells, so getting $d\,n^d$ distinct basis elements of $C_{d-1}$. (b) Then we complete this bases by adjoining $n^{d-1}$ boundary elements for each of the $d$ dimensions of the image, so providing further $d\,n^{d-1}$ basis elements for $C_{d-1}$. The dimension of $C_{d-1}$, and therefore the number of rows of $[\partial_d]$ matrix is $d\,(n^{d-1}+n^{d}) = d\,n\,(1+n)^{d-1}$. The number of column equates the number of basis elements of $C_d$, i.e.~the number $n^d$ of brick elements.

@@

As we have seen, we have $2d$ non-zero elements for each column of $[\partial_d]$, so that their total number is $2d\,n^d$. The number of matrix element is $d\,n\,(1+n)^{d-1} \times n^d$, giving a ratio of 
\[
\frac{\mbox{non-zero\ elements}}{\mbox{total\ elements}} = 
\frac{2d\times n^d}{d\,n\,(1+n)^{d-1} \times n^d} =
\frac{2}{n+n^d}
\]
Using sparse matrices in CSC (Compressed Sparse Column) format we get a storage size:
\[
mem([\partial_d]_{n^d}) = 2\times \#\mbox{nzero} + \#\mbox{columns} = 2\times 2d\,n^d + n^d = (4d+1)n^d.
\]
In conclusion, for brick size $n=64$, the matrix $[\partial_d]$ requires for 2D images $9\times 64^2=36,864$ memory elements, and for 3D images $13\times 64^3=3,407,872$ memory elements. Counting the bytes for the standard implementation of a sparse binary matrix (1 byte for values and 8 bytes for indices) we get $(18d+8)n^d$ bytes, giving $176$\,KB for 2D and $15.872$\,MB for 3D.

% \begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=0.4\linewidth]{figs/grid.pdf} 
%   \caption{Faces on the brick boundary}
%   \label{fig:brickboundary}
% \end{figure}

\subsection{Brick boundary mapping}\label{sec:brick-mapping}

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. Our aim is to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the brick, in order to decompose the input $Image(u,v,w)$ into a fair number of bricks
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n} \simeq \frac{uvw}{n^3}.
\] 
Then, we consider each segment portion $c_{i,j,k} = S\cap \B(i,j,k,n)$ and compute its local coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then, assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$, with $d=3$. Finally, compute a matrix $\T{\textbf{B}}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{\textbf{B}} = [\partial_3(n)]\, \T{S}.
\]
where the boundary matrix has dimension $n^d \times dn(n+1)^{d-1}$.
Of course, the $\T{\textbf{B}}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S \cap \B(i,j,k,n)$, and the number of matrix rows is the dimension $n^d$ of the linear space $C_2$ of 2-chains.

@@

\subsection{Brick-level parallelism}\label{sec:brick-parallelism}
In the computational pipeline introduced in this paper, several steps can be efficiently performed in parallel at the image-brick level, depending on the embarrassingly data-parallel nature of the problem. In particular, little effort is needed to separate the problem into a several of parallel tasks $S_{i,j,k}$, using multiarray slicing. The granularity of parallelism, depending on the brick size $n$, is further enforced by the computation of a single boundary matrix $[\partial_d(n)]$ depending on $n$, so that the initial communication cost of broadcasting the matrix to nodes can be carefully controlled, and finely tuned depending on the system architecture. The whole approach is appropriate  for SIMD (Single Instruction, Multiple Data) hybrid architectures of CPUs and GPUs, since only the initial brick setup of boundary matrix and image slices, as well the final collection of computed surface portions, require inter-process communication.

### 2020-04-26 src/examples.tex
submission

% \includegraphics[height=0.3\textwidth]{figs/nrn10_100_green.png} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{src/figs/ircad01_segmentation_65.png}
\includegraphics[width=0.45\textwidth]{src/figs/ircad01_liver_tricolore_01.png}
\includegraphics[width=0.45\textwidth]{src/figs/ircad01_porta_blue_01.png}
\includegraphics[width=0.45\textwidth]{src/figs/nrn10_200_pink_02.png}
% \includegraphics[width=0.4\textwidth]{figs/liver_01_red_3.png} 
% % \vspace{0.05\textwidth}
% \includegraphics[width=0.4\textwidth]{figs/portalvein_01_yellow_3.png} 
% % \vspace{0.01\textwidth}
% \includegraphics[height=0.4\textwidth]{src/figs/nrn10_100_magenta_high_res.png} 
% % \includegraphics[height=0.4\textwidth]{figs/nrn10_100_low_res.png} 
% % \vspace{0.05\textwidth}
% \includegraphics[width=0.4\textwidth]{figs/porta_smoothing_2.png} 
\caption{
Liver structures. The upper left image shows one slice  with segmented organs from  the Ircad dataset \cite{ircad}.
The other images show the triangulated isosurfaces of the macroscopic and microscopic strucures of the liver extracted with \textsc{lar-surf} algorithm. 
On the upper right image can be seen the wire frame model of the human liver along with portal vein and hepatic artery. 
% Both right images show the Portal Vein from the same Ircad dataset \cite{ircad} 
The left bottom image shows the detail of the portal vein  
with resolution $1.6\times0.57\times0.57$ $[mm]$.
On the bottom right image is shown the microvasculature of a pig liver based on corrosion cast prepared by Eberlova
\cite{eberlova2017use}. The size of the speciment is 0.936 $[mm]$ along each axis and the resolution of the Micro-CT data is 4.682 $\mu{}m$.
} \label{fig:example_liver_macro_micro}
\end{figure}

@@

The implementation of our algorithm is available in LarSurf package. 
Use of this package can be seen on listings \ref{lst:example1} where the liver segmentation with 2865131 voxels from
Ircad dataset is used as an input for our surface extraction algorithm. The size of 3D volumetric 
image is $129 \times 512 \times 512$
and the voxel resolution is $1.6\times0.57\times0.57$ [mm]. 
The output surface model created by  182124 triangles and the number of vertices is 90822. The visualization can 
be seen on figure \ref{fig:example_liver_macro_micro}. 
The portal vein surface extraction can be performed with a small change of input path in code.
The The 3D image resolution is the same. 
It can be seen in the same figure. 
The number of input voxels is 103533. 
The output surface is created by 90822 vertices and 182124 triangles.

@@

The right image of figure \ref{fig:example_liver_macro_micro} is surface of a microvasculature of pig liver. 
Volumetric image is based on Micro-CT data of corrosion casts of pig liver.
\cite{eberlova2017use}.
The size of the visualized data is $100\times100\times100$ voxels and size of voxel is 4.682 $\mu{}m$.
The number of of triangles is 
544784 and number of vertices is 272826.

>>>>

% \includegraphics[height=0.3\textwidth]{figs/nrn10_100_green.png} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\begin{figure}
\centering
\includegraphics[width=0.40\textwidth]{src/figs/ircad01_segmentation_65.png}
\includegraphics[width=0.40\textwidth]{src/figs/ircad01_liver_tricolore_01.png}
\includegraphics[width=0.40\textwidth]{src/figs/ircad01_porta_blue_01.png}
\includegraphics[width=0.40\textwidth]{src/figs/nrn10_200_pink_02.png}
% \includegraphics[width=0.4\textwidth]{figs/liver_01_red_3.png} 
% % \vspace{0.05\textwidth}
% \includegraphics[width=0.4\textwidth]{figs/portalvein_01_yellow_3.png} 
% % \vspace{0.01\textwidth}
% \includegraphics[height=0.4\textwidth]{src/figs/nrn10_100_magenta_high_res.png} 
% % \includegraphics[height=0.4\textwidth]{figs/nrn10_100_low_res.png} 
% % \vspace{0.05\textwidth}
% \includegraphics[width=0.4\textwidth]{figs/porta_smoothing_2.png} 
\caption{
Liver structures. The upper left image shows one slice  with segmented organs from  the Ircad dataset \cite{ircadb}.
The other images show the triangulated isosurfaces of the macroscopic and microscopic structures of the liver extracted with \textsc{lar-surf} algorithm. 
The wireframe model of the human liver along with the portal vein and hepatic artery can be seen on the upper right image. 
The left bottom image shows the detail of the portal vein  
with resolution $1.6\times0.57\times0.57$ $[mm]$.
On the bottom right image is shown the microvasculature of a pig liver based on corrosion cast prepared by Eberlova
\cite{eberlova2017use}. The size of the specimen is 0.936 $[mm]$ along each axis and the resolution of the Micro-CT data is 4.682 $\mu{}m$.
} \label{fig:example_liver_macro_micro}
\end{figure}
% Both right images show the Portal Vein from the same Ircad dataset \cite{ircad}

@@

% The implementation of our algorithm is available in LarSurf package. 
Use of the \texttt{LarSurf.jl} package can be seen on listings \ref{lst:example1} where the liver segmentation with 2865131 voxels from the
Ircad dataset is used as an input for our surface extraction algorithm. The size of 3D volumetric 
image is $129 \times 512 \times 512$
and the voxel resolution is $1.6\times0.57\times0.57$ [mm]. 
The output liver surface model is formed by  182124 triangles and the number of vertices is 90822. 
The visualization can 
be seen in Fig. \ref{fig:example_liver_macro_micro}.

@@

The portal vein surface extraction can be performed with a small change of input path in code.
The 3D image resolution is the same. 
% It can be seen in the same figure. 
The number of input voxels is 103533. 
The output surface is created by 90822 vertices and 182124 triangles.

The right bottom image of Fig. \ref{fig:example_liver_macro_micro} is the surface of the microvasculature of pig liver. 
The volumetric image is based on Micro-CT data of corrosion casts of pig liver.
\cite{eberlova2017use}.
The size of the visualized data is $100\times100\times100$ voxels and the size of the voxel is 4.682 $\mu{}m$.
The number of triangles is 
544784 and the number of vertices is 272826.
